No Dogs On The Couch$_{Base}$

### 2012-11-28 Assignment6/paper/paper.tex
Consolidated data into 3 different tables, moved JSON and ER-diagram to be an appendix at the end of the paper, changed things in the tables so that they would fit into a single column.
If someone thinks more description is needed for a table try to put it into the
caption so that the tables themselves fit into a single column (it's a bit
cleaner this way since the numbers we have are not very big, it's just the
names of the columns, etc.)

Also, added the runtime for the assign task workload in couch.log

% Paper for the ChouchBase assignment.

\documentclass[]{IEEEtran}

\usepackage{float}
\usepackage{url}
\usepackage{graphicx}
\usepackage{color}

@@

This paper primarily focuses on considerations for \textit{CGAT}'s database
architecture. These considerations are made with regards to a MySQL cluster
versus a Couchbase cluster to determine which database architecture is most
efficient for running \textit{CGAT}. Section \ref{sec:motive} discusses our
motivation to compare Couchbase and MySQL in this paper, while section
\ref{sec:design} talks about how we designed experimental workloads to use for
our comparison of Couchbase vs MySQL. In section \ref{sec:workload} we
describe the workloads and how they are relevant to \textit{CGAT}, some brief
implementation details are outlined in section \ref{sec:implementation}, we
evaluate our results in section \ref{sec:eval} and finally conclude in
section \ref{sec:conclusions}.

@@

\begin{figure}
	\includegraphics[width=5.5in]{SQLDiagram.eps}
	\caption{The SQL diagram for CGAT.}
	\label{fig:SQLDiagram}
\end{figure}

%\clearpage

\begin{figure}%
	\includegraphics{user.eps}
	\caption{The JSON for a user.}
	\label{fig:JSON-user}
\end{figure}%

\begin{figure}%
	\includegraphics[height=70mm]{contig.eps}
         \caption{The JSON for a contig.}
         \label{fig:JSON-contig}
\end{figure}%

\begin{figure}%
	\includegraphics[height=100mm]{annotation.eps}
	\caption{The JSON for an annotation.}
	\label{fig:JSON-annotation}
\end{figure}%

\begin{figure}%
	\includegraphics[height=20mm]{group.eps}
	\caption{The JSON for a group.}
	\label{fig:JSON-group}
\end{figure}%

\clearpage

@@

\subsection{Read/Write}
This is a test workload to examine the performance of writing and reading from
the same object in succession. This is a workload contrived purely for better
comparison of Couchbase vs MySQL instead of workloads representative of what
\textit{CGAT} will rely on. This is a naive, balanced workload.

@@

@@

\subsection{Assigning a Task}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
	{\bf Metric} & {\bf Result} \\
	\hline
        \# of runs & 100000 \\
        \hline
        MySQL runtime (ms) & 39325189 \\
	\hline
        Couchbase runtime (ms) & ?* \\
        \hline
\end{tabular}
\end{center}

From the results it is obvious that MySQL performs much better than Couchbase. The result we have for
Couchbase is actually a minimum runtime. It was running so long we killed it to get more results. Even with this
in mind, MySQL vastly outperforms Couchbase. We believe this is because during the process of assigning a task
a couple values have to be read into memory before a finally write can be done. Compared to MySQL where its a simple 
ad-hoc query to find the necessary values prior to inserting the new task.

@@

\subsection{Profiling}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
	{\bf Metric} & {\bf Result} \\
	\hline
        \# of runs & 100000 \\
        \hline
        MySQL runtime (ms) & 2908257 \\
	\hline
        Couchbase runtime (ms) & 7008818 \\
        \hline
\end{tabular}
\end{center}

Like the last workload, MySQL outperforms Couchbase. We believe that this is because this workload is read-heavy so several
full documents are retrieved per run.  Again, this workload could be improved in Couchbase with the use of unstructured queries.

\subsection{Publishing}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
	{\bf Metric} & {\bf Result} \\
	\hline
        \# of runs & 100000 \\
        \hline
        MySQL runtime (ms) & 510837 \\
	\hline
        Couchbase runtime (ms) & 1830064 \\
        \hline
\end{tabular}
\end{center}

Again, a similar pattern in the results as the last two workloads. This was another workload that required some queries prior to inserting
a new annotation. We believe this is another case where unstructured queries could vastly improve performance.

\subsection{Adding and Removing Group Membership}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
	{\bf Metric} & {\bf Result} \\
	\hline
        \# of runs & 100000 \\
        \hline
        MySQL runtime (ms) & 316673 \\
	\hline
        Couchbase runtime (ms) & 9086304 \\
        \hline
\end{tabular}
\end{center}

To our surprise this workload performed extremely bad on Couchbase. We were not expecting this because this is a 50\% read
and 50\% write workload. Having to do lots of reads prior to inserting new group memberships really hurt Couchbase. This is most definitely 
a case where having unstructured queries or document support would greatly improve performance. The way we store group memberships
is as a sub-field in the user and group objects. By only queries for those fields we could reduce the amount of data we return to the workload harness by over 70\%.

\subsection{Uploading Contigs}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
	{\bf Metric} & {\bf Result} \\
	\hline
        \# of runs & 10000 \\
        \hline
        MySQL runtime (ms) & 119892 \\
	\hline
        Couchbase runtime (ms) & 20133 \\
        \hline
\end{tabular}
\end{center}

The results from this test were no surprise. Couchbase demolished MySQL when most of the workload was insertions. Even with
contig objects containing strings of over 50k characters Couchbase performed amazingly well. By not having to lock key-value pairs
Couchbase can get some awesome write performance.

\subsection{Read and Writing Performance}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
	{\bf Metric} & {\bf Result} \\
	\hline
        \# of runs & 100000 \\
        \hline
        MySQL runtime (ms) & 103808 \\
	\hline
        Couchbase runtime (ms) & 58581 \\
        \hline
\end{tabular}
\end{center}

The results from this test were not that surprising based on the results from our other tests. It is obvious that Couchbase is much faster in terms of writes plus Couchbase can leverage Memcache in this workload to speed up queries to the database. For this workload we decided to gather a bunch more numbers to see what the minimum, maximum, mean, median and standard deviation of the individual reads and writes for each system. Below we list the six tables and discuss what the results mean.

\begin{center}
\begin{tabular}{|l|l|}
	\hline
  	\multicolumn{2}{|c|}{{\bf SQL read after writes of same object}} \\
 	\hline
  	 Min (ms) & 0 \\
  	\hline
         Max (ms) & 25 \\
  	\hline
  	Mean (ms) & 0.473 \\
 	\hline
  	Median (ms) & 0 \\
  	\hline
         Standard Deviation (ms) & 0.685 \\
  	\hline
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
  	\multicolumn{2}{|c|}{{\bf SQL write after reads of same object}} \\
 	\hline
  	 Min (ms) & 0 \\
  	\hline
         Max (ms) & 70 \\
  	\hline
  	Mean (ms) & 0.558 \\
 	\hline
  	Median (ms) & 1 \\
  	\hline
         Standard Deviation (ms) & 0.859 \\
  	\hline
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
  	\multicolumn{2}{|c|}{{\bf SQL combined results}} \\
 	\hline
  	 Min (ms) & 0 \\
  	\hline
         Max (ms) & 70 \\
  	\hline
  	Mean (ms) & 0.516 \\
 	\hline
  	Median (ms) & 0 \\
  	\hline
         Standard Deviation (ms) & 0.778 \\
  	\hline
\end{tabular}
\end{center}

@@

\begin{center}
\begin{tabular}{|l|l|}
	\hline
  	\multicolumn{2}{|c|}{{\bf Couchbase read after writes of same object}} \\
 	\hline
  	 Min (ms) & 0 \\
  	\hline
         Max (ms) & 71 \\
  	\hline
  	Mean (ms) & 0.562 \\
 	\hline
  	Median (ms) & 1 \\
  	\hline
         Standard Deviation (ms) & 0.859 \\
  	\hline
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
  	\multicolumn{2}{|c|}{{\bf Couchbase write after reads of same object}} \\
 	\hline
  	 Min (ms) & 0 \\
  	\hline
         Max (ms) & 48 \\
  	\hline
  	Mean (ms) & 0.017 \\
 	\hline
  	Median (ms) & 0 \\
  	\hline
         Standard Deviation (ms) & 0.213 \\
  	\hline
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{|l|l|}
	\hline
  	\multicolumn{2}{|c|}{{\bf Couchbase combined results}} \\
 	\hline
  	 Min (ms) & 0 \\
  	\hline
         Max (ms) & 71 \\
  	\hline
  	Mean (ms) & 0.29 \\
 	\hline
  	Median (ms) & 0 \\
  	\hline
         Standard Deviation (ms) & 0.682 \\
  	\hline
\end{tabular}
\end{center}

@@

>>>>

% Paper for the ChouchBase assignment.
\documentclass[10pt, conference, compsocconf]{IEEEtran}

\usepackage{float}
%\usepackage{url}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

@@

This paper primarily focuses on considerations for \textit{CGAT}s database
architecture. These considerations are made with regards to a MySQL cluster
versus a Couchbase cluster to determine which database architecture is most
efficient for running \textit{CGAT}. Section \ref{sec:motive} discusses our
motivation to compare Couchbase and MySQL in this paper, while section
\ref{sec:design} talks about how we designed experimental workloads to use for
our comparison of Couchbase vs MySQL. In section \ref{sec:workload} we
describe the workloads and how they are relevant to \textit{CGAT}, some brief
implementation details are outlined in section \ref{sec:implementation}, we
evaluate our results in section \ref{sec:eval} and finally conclude in
section \ref{sec:conclusions}.

@@

\begin{table}[t]
   \centering
      \begin{tabular}{| l | c | c | c |}
         \hline
            {\textbf{Workload}} &
            {\textbf{MySQL}} &
            {\textbf{Couchbase}} \\
         \hline
            Task Assignment         & 39325189ms & 157431049ms \\
         \hline
            Profile View            & 2908257ms & 7008818ms \\
         \hline
            Annotation Publication  & 510837ms & 1830064ms \\
         \hline
            Group Modification      & 316673ms & 9086304ms \\
         \hline
            Contig Upload           & 119892ms & 20133ms \\
         \hline
            Read/Write              & 103808ms & 58581ms \\
         \hline
      \end{tabular}
      \caption{Comparison of runtimes in ms for workloads using MySQL and Couchbase.
               Each workload was run for 100000 iterations.}
      \label{tab:workload_perf}
\end{table}

@@

\subsection{Read/Write Performance}
This is a test workload to examine the performance of writing and reading from
the same object in succession. This is a workload contrived purely for better
comparison of Couchbase vs MySQL instead of workloads representative of what
\textit{CGAT} will rely on. This is a naive, balanced workload.

@@

\begin{table}[t]
   \centering
      \begin{tabular}{| l | c | c | c |}
         \hline
            \multicolumn{4}{|c|}{{\textbf{MySQL read/write performance}}} \\
         \hline
            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
         \hline
            Min (ms) & 0 & 0 & 0 \\
         \hline
            Max (ms) & 25 & 70 & 70 \\
         \hline
            Mean (ms) & 0.473 & 0.558 & 0.516 \\
         \hline
            Median (ms) & 0 & 1 & 0 \\
         \hline
            Standard Deviation (ms) & 0.685 & 0.859 & 0.778 \\
         \hline
      \end{tabular}
      \caption{Comparison of reads and writes to the same object in MySQL.}
      \label{tab:sql_perf}
\end{table}

\begin{table}[t]
   \centering
      \begin{tabular}{| l | c | c | c |}
         \hline
            \multicolumn{4}{|c|}{{\textbf{Couchbase read/write performance}}} \\
         \hline
            & \textbf{RAW} & \textbf{WAR} & \textbf{combined} \\
         \hline
            Min (ms) & 0 & 0 & 0 \\
         \hline
            Max (ms) & 71 & 48 & 71 \\
         \hline
            Mean (ms) & 0.562 & 0.017 & 0.290 \\
         \hline
            Median (ms) & 1 & 0 & 0 \\
         \hline
            Standard Deviation (ms) & 0.859 & 0.213 & 0.682 \\
         \hline
      \end{tabular}
      \caption{Comparison of reads and writes to the same object in couchbase}
      \label{tab:couch_perf}
\end{table}

@@

\subsection{Task Assignment}
From the results it is obvious that MySQL performs much better than Couchbase.
The result we have for Couchbase is actually a minimum runtime. It was running
so long we killed it to get more results. Even with this in mind, MySQL vastly
outperforms Couchbase. We believe this is because during the process of
assigning a task a couple values have to be read into memory before a finally
write can be done. Compared to MySQL where its a simple ad-hoc query to find
the necessary values prior to inserting the new task.

@@

\subsection{Profile View}
Like the last workload, MySQL outperforms Couchbase. We believe that this is
because this workload is read-heavy so several full documents are retrieved per
run.  Again, this workload could be improved in Couchbase with the use of
unstructured queries.

\subsection{Annotation Publication}
Again, a similar pattern in the results as the last two workloads. This was
another workload that required some queries prior to inserting a new
annotation. We believe this is another case where unstructured queries could
vastly improve performance.

\subsection{Group Modification}
To our surprise this workload performed extremely bad on Couchbase. We were not
expecting this because this is a 50\% read and 50\% write workload. Having to
do lots of reads prior to inserting new group memberships really hurt
Couchbase. This is most definitely a case where having unstructured queries or
document support would greatly improve performance. The way we store group
memberships is as a sub-field in the user and group objects. By only queries
for those fields we could reduce the amount of data we return to the workload
harness by over 70\%.

\subsection{Contig Uploads}
The results from this test were no surprise. Couchbase demolished MySQL when
most of the workload was insertions. Even with contig objects containing
strings of over 50k characters Couchbase performed amazingly well. By not
having to lock key-value pairs Couchbase can get some awesome write
performance.

\subsection{Read/Write Performance}
The results from this test were not that surprising based on the results from
our other tests. It is obvious that Couchbase is much faster in terms of writes
plus Couchbase can leverage Memcache in this workload to speed up queries to
the database. For this workload we decided to gather a bunch more numbers to
see what the minimum, maximum, mean, median and standard deviation of the
individual reads and writes for each system. Below we list the six tables and
discuss what the results mean.

@@

@@

\newpage
\appendices
\begin{figure*}
   \centering
   \includegraphics[height=90mm]{SQLDiagram.eps}
	\caption{The SQL diagram for CGAT.}
	\label{fig:SQLDiagram}
\end{figure*}

\begin{figure*}{t}
   \centering
   \subfloat[Contig JSON]{\label{fig:JSON-contig}\includegraphics[height=70mm]{contig.eps}}
   \subfloat[Annotation JSON]{\label{fig:JSON-annotation}\includegraphics[height=80mm]{annotation.eps}}
\end{figure*}

\begin{figure*}{t}
   \centering
   \subfloat[User JSON]{\label{fig:JSON-user}\includegraphics[height=90mm]{user.eps}}
   \subfloat[Group JSON]{\label{fig:JSON-group}\includegraphics[height=25mm]{group.eps}}
\end{figure*}

%\begin{figure*}
%  \includegraphics[height=90mm]{user.eps}
%	\caption{The JSON for a user.}
%	\label{fig:JSON-user}
%\end{figure*}
%\begin{figure*}
%  \includegraphics[height=25mm]{group.eps}
%	\caption{The JSON for a group.}
%	\label{fig:JSON-group}
%\end{figure*}
%\begin{figure*}
%  \includegraphics[height=80mm]{contig.eps}
%  \caption{The JSON for a contig.}
%  \label{fig:JSON-contig}
%\end{figure*}
%\begin{figure*}
%  \includegraphics[height=90mm]{annotation.eps}
%	\caption{The JSON for an annotation.}
%	\label{fig:JSON-annotation}
%\end{figure*}
